---
title: ggplot2 - Axis & Plot Labels
date: '2017-03-05'
slug: axis-plot-labels
topics:
  - data-visualization
tags:
  - ggplot2
---

### Introduction

This is the fifth post in the series **Elegant Data Visualization with ggplot2**. In the previous [post](https://blog.aravindhebbali.com/), we learnt about aesthetics. In this post, we will learn to:

- add title and subtitle to the plot
- modify axis labels
- modify axis range
- remove axis
- format axis

```{r axis1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

#### Title & Subtitle

Let us create a simple scatter plot with which we will work in the rest of the post.

```{r axis2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg))
```

There are two ways to add title to a plot:

- `ggtitle()`
- `labs()`

#### ggtitle()

Let us explore the `ggtitle()` function first. It takes two arguments:

- label: title of the plot
- subtitle: subtitle of the plot

##### Title & Subtitle

```{r axis3, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  ggtitle(label = 'Displacement vs Mileage', subtitle = 'disp vs mpg')
```

#### Axis Labels

You can add labels to the axis using:

- `xlab()`
- `ylab()`
- `labs()`

Let us add labels using `xlab()` for the X axis and `ylab()` for the Y axis.

```{r axis4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  xlab('Displacement') + ylab('Miles Per Gallon')
```

#### Labs

The `labs()` function can be used to add the following to a plot:

- title
- subtitle
- X axis label
- Y axis label

Let us add title and axis labels using `labs()`.

```{r axis5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  labs(title = 'Displacement vs Mileage', subtitle = 'disp vs mpg', 
    x = 'Displacement', y = 'Miles Per Gallon')
```

#### Axis Range

In certain scenarios, you may want to modify the range of the axis. In ggplot2, we can achieve this using:

- `xlim()`
- `ylim()`
- `expand_limits()`

`xlim()` and `ylim()` take a numeric vector of length 2 as input. `expand_limits()` takes two numeric vectors (each of length 2), one for each axis. In all of the above functions, the first element represents the lower limit and the second element represents the upper limit.

Let us modify the range of the X axis using `xlim()`. The lower limit will be 0 and the upper limit 600.

```{r axis6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  xlim(c(0, 600))
```

The range of the Y axis will be modified using `ylim()`. The lower limit will be 0 and the upper limit will be 40.

```{r axis7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  ylim(c(0, 40))
```

Instead of using `xlim()` and `ylim()`, we can use `expand_limits()` to modify the range of both the X and Y axis. 

```{r axis8, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  expand_limits(x = c(0, 600), y = c(0, 40))
```

#### Remove Axis Labels

If you want to remove the axis labels all together, use the `theme()` function. It allows us to modify every aspect of the theme of the plot. Within `theme()`, set the following to `element_blank()`.

- `axis.title.x`
- `axis.title.y`

In the below example, we use `theme()` to remove the X and Y axis labels.

```{r axis9, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank())
```

#### Format Title & Axis Labels

To format the title or the axis labels, we have to modify the theme of the plot using the `theme()` function. We can modify: 

- color
- font family
- font face
- font size 
- horizontal alignment 
- and angle

`element_text()` is used to modify the properties of text. In the below examples, we use it 
to modify the plot title (which is a text).

##### Color

```{r axis10, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(color = 'blue'))
```

##### Font Family

```{r axis11, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(family = 'Arial'))
```

##### Font Face

The font face can be any of the following:

- `plain` 
- `bold`
- `italic` 
- `bold.italic`

##### Font Face 

```{r axis12, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(face = 'bold'))
```

##### Font Size

```{r axis13, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(size = 8))
```

##### Horizontal Alignment

```{r axis14, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(hjust = 0.3))
```

Let us use `element_text()` and `theme()` to modify the appearance of
the title and axis labels.

#### Title

```{r axis15, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ggtitle('Diaplacement vs Mileage') +
  theme(plot.title = element_text(color = 'blue', family = 'Arial',
    face = 'bold', size = 12, hjust = 0.5))
```

#### X Axis Label

```{r axis16, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + xlab('Diaplacement') +
  theme(axis.title.x = element_text(color = 'blue', family = 'Arial',
    face = 'bold', size = 8, hjust = 0.5, angle = 15))
```

#### Y Axis Label

```{r axis17, fig.align='center', message=FALSE, warning=FALSE}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) + ylab('Miles Per Gallon') +
  theme(axis.title.y = element_text(color = 'blue', family = 'Arial',
    face = 'italic', size = 8, vjust = 0.3, angle = 90))
```

### Summary

In this post, we learnt to:

- add title and subtitle to the plot
- modify axis labels
- modify axis range
- remove axis
- format axis

### Up Next..

In the next [post](https://blog.aravindhebbali.com), we will learn to add text annotations to plots.

