---
title: Importing Data into R - Part 3
author: Aravind Hebbali
date: '2017-12-25'
slug: importing-data-into-r-part-3
categories:
  - data wrangling
tags:
  - import
  - xml
  - json
---

## Introduction

This is the third post in the series **Importing Data into R**. You can find the link to the previous posts below:

- [Importing Data into R - Part 1](https://blog.aravindhebbali.com/2017/12/24/a-complete-guide-to-importing-data-into-r-part-1/)
- [Importing Data into R - Part 2](https://blog.aravindhebbali.com/2017/12/25/importing-data-into-r-part-2/)

In the previous post, we explored reading data from excel spreadsheets and files from other statistical softwares such as SAS, SPSS and STATA. In this post, we will learn to read data from:

- JSON
- and XML

## Libraries, Data & Code

We will use the [xml2](https://cran.r-project.org/package=xml2) and [XML](https://cran.r-project.org/package=XML) package for reading data from XML files and [jsonlite](https://cran.r-project.org/package=jsonlite) package to read data from JSON files. The data sets can be downloaded from [here](https://github.com/rsquaredacademy/datasets) and the codes from [here](https://gist.github.com/aravindhebbali/65dc2af57b8eda098f2b878095c5aaa4).

```{r lib, message=FALSE}
library(jsonlite)
library(xml2)
library(XML)
library(tibble)
library(purrr)
library(magrittr)
library(stringr)
```

## Read Data

The first step is to read XML data into R using `read_xml()`. Let us
read the data from the **books.xml** file.

```{r imp15}
books <- read_xml('books.xml')
```

## Root Node

Let us look at the root node of the xml file using `xml_root()`.

```{r imp16}
xml_root(books)
```

`xml_name()` returns the name of an node. Here it returns the name of the 
root node. It can also be used to modify the name of a node.

```{r imp17}
xml_name(books)
```

We can view the contents using `xml_contents()`.

```{r imp18}
xml_contents(books)
```

`xml_length()` will return the number of child nodes.

```{r imp19}
xml_length(books)
```

## Child Node

Returns nodeset but only the elements and not their contents.

```{r imp20, collapse = FALSE}
xml_children(books) 

books %>% 
  xml_children() %>% 
  xml_contents()

books %>% 
  xml_children() %>% 
  xml_length()
```

View the contents of a particular child node. Specify the node
using `search`.

```{r imp21, collapse = FALSE}
xml_child(books)

xml_child(books, search = 2)
```

Use `xml_parent()` to view the parent node of any child node.

```{r imp22}
books %>%
    xml_child(search = 2) %>%
    xml_parent()
```

## Attributes

We can extract attributes using `xml_attrs()`. If we want to extract a specific
attribute value, use `xml_attr()` and specify the attribute name. In the books data,
`book id` is an attribute. 

```{r imp23, collapse = FALSE}
xml_find_all(books, './/book') %>% 
    map_chr(xml_attrs)

xml_find_all(books, './/book') %>% 
  xml_attr(attr = "id")
```

Since we want to convert the xml data into a tibble, let us use
`xml_find_all()` to extract data for each of the columns. We will
use the node name prepended by `.//` to extract the data.

```{r imp24}
xml_find_all(books, './/author')
```

We want the author names and not the nodes. Let us extract the author names
as string using `xml_text()`. In case of numbers, we can use `xml_double()`.

```{r imp25}
books %>%
  xml_find_all(xpath = './/author') %>%
  xml_text()
```

`xml_find_all()` will find all the matches. If you want to find a specific match use
`xml_find_first()`.

```{r imp26, collapse = FALSE}
xml_find_first(books, './/author')

books %>%
  xml_find_first(xpath = './/author') %>%
  xml_text()
```

#### Author Name

```{r imp27}
books %>%
  xml_find_all(xpath = './/author') %>%
  xml_text()
```

#### Book Title

```{r imp28}
books %>%
  xml_find_all(xpath = './/title') %>%
  xml_text()
```

#### Book Genre

```{r imp29}
books %>%
  xml_find_all(xpath = './/genre') %>%
  xml_text()
```

#### Book Price

Since book price is numeric, we will use `xml_double()` to extract the
data as number and not text.

```{r imp30, collapse = FALSE}
books %>%
    xml_find_all(xpath = './/price')

books %>%
    xml_find_all(xpath = './/price') %>%
    xml_double()
```

#### Publishing Date

```{r imp31}
books %>%
  xml_find_all(xpath = './/publish_date') %>%
  xml_text()
```

#### Book Description

```{r imp32}
books %>%
  xml_find_all(xpath = './/description') %>%
  xml_text() %>%
  str_replace_all("[\n]", "") %>%
  str_replace_all(pattern = "      ", " ") %>%
  str_replace_all(pattern = "  ", " ")
```

## Book Data

Let us create a tibble to store the xml data.

```{r imp33, collapse = FALSE}
book <- 
  tibble(
    author = books %>%
      xml_find_all(xpath = './/author') %>% 
      xml_text(),
    title = books %>%
      xml_find_all(xpath = './/title') %>%
      xml_text(),
    genre = books %>%   
      xml_find_all(xpath = './/genre') %>%
      xml_text(),
    price = books %>%
      xml_find_all(xpath = './/price') %>%
      xml_double(),
    published = books %>%
      xml_find_all(xpath = './/publish_date') %>%
      xml_text(),
    description = books %>%
      xml_find_all(xpath = './/description') %>%
      xml_text() %>%
      str_replace_all("[\r\n]", "") %>%
      str_replace_all(pattern = "      ", " ") %>%
      str_replace_all(pattern = "  ", " "))

book
```


