---
title: SQL for Data Science - Part 1
date: '2017-12-25'
slug: sql-for-data-science-part-1
categories:
  - database
tags:
  - sql
---

```{r setup, eval=TRUE, message=FALSE, echo=FALSE}
library(dplyr)
library(DBI)
library(RSQLite)
knitr::opts_chunk$set(R.options=list(max.print=50))
```

### Introduction

This is the third post in the series **R & Databases**. You can find the links to the other two posts of this series below:

- [Quick Guide: R & SQLite](https://rsquaredacademy.github.io/blog/post/quick-guide-r-sqlite/)
- [Data Wrangling with dbplyr](https://rsquaredacademy.github.io/blog/post/data-wrangling-with-dbplyr)
- [SQL for Data Science - Part 2](https://rsquaredacademy.github.io/blog/post/sql-for-data-science-part-2)

In this post, we will learn to:

### Libraries, Code & Data

We will use the following libraries in this post:

- [DBI](http://readr.tidyverse.org/)
- [RSQLite](https://rstats-db.github.io/RSQLite/)
- [dbplyr](http://dbplyr.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) and code can be downloaded from [here](https://gist.github.com/rsquaredacademy/f44774d9c6e115c08d5e2e693380a00f).

### Set Up

```{r setup2, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
copy_to(con, ecom)
```

#### Select Single Column

```{r sql1}
dbGetQuery(con, "select device from ecom")
```

#### Select Multiple Columns

```{r sql2}
dbGetQuery(con, "select referrer, device, purchase  from ecom")
```

#### Select All Columns

```{r sql3}
dbGetQuery(con, "select * from ecom")
```

#### Select All Columns (Limit)

```{r sql4}
dbGetQuery(con, "select * from ecom limit 10")
```

#### Select Distinct

```{r sql5}
dbGetQuery(con, "select distinct referrer from ecom")
```

#### Count

```{r sql6}
dbGetQuery(con, "select count(*) from ecom")
```

### Filter

You now know how to select columns and perform basic counts. This chapter will focus on filtering your results.

In SQL, the WHERE keyword allows you to filter based on both text and numeric values in a table. There are a few different comparison operators you can use:

- `=` equal
- `<>` not equal
- `<` less than
- `>` greater than
- `<=` less than or equal to
- `>=` greater than or equal to


#### Filter

```{r sql15}
dbGetQuery(con, "SELECT * FROM ecom WHERE duration > 300")
```

#### Filter Text

```{r sql7}
dbGetQuery(con, "SELECT * FROM ecom WHERE device == 'mobile'")
```

#### Where And

```{r sql8}
dbGetQuery(con, "SELECT * FROM ecom WHERE n_visit > 3 AND duration > 100")
```

#### Where And Or - Part 1

#### Where And Or - Part 2

```{r sql9}
dbGetQuery(con, "SELECT * FROM ecom WHERE (n_visit == 5 OR n_visit == 3) 
  AND (device = 'mobile' OR device = 'tablet')")
```

#### Between - Part 2

```{r sql10}
dbGetQuery(con, "SELECT * FROM ecom WHERE n_visit BETWEEN 1 AND 3
  AND device = 'mobile'")
```

#### Where In

```{r sql11}
dbGetQuery(con, "SELECT * FROM ecom WHERE n_visit IN (2, 4, 6, 8, 10)")
```

#### Introduction to NULL and IS NULL

```{r sql12}
dbGetQuery(con, "SELECT * FROM ecom WHERE device IS NULL")
```

#### LIKE and NOT LIKE - Part 1

```{r sql13}
dbGetQuery(con, "SELECT * FROM ecom WHERE country LIKE 'P%'")
```

#### LIKE and NOT LIKE - Part 3

```{r sql14}
dbGetQuery(con, "SELECT * FROM ecom WHERE country LIKE '_o%'")
```

### Summary

### Up Next..

In the next [post](https://rsquaredacademy.github.io/blog/post/sql-for-data-science-part-2), we will learn advanced SQL commands.


