---
title: Data Wrangling with dplyr - Part 3
author: Aravind Hebbali
date: '2017-12-25'
slug: data-wrangling-with-dplyr-part-3
categories:
  - data wrangling
tags:
  - dplyr
---

## Introduction

In the previous [post](), we learnt to combine tables using different `*_join` functions provided in dplyr. In this module, we will explore a set of helper functions in order to:

- extract unique rows 
- rename columns
- sample data
- extract columns
- slice rows
- arrange rows
- compare tables
- extract/mutate data using predicate functions
- count observations for different levels of a variable

<br>

## Libraries, Code & Data

We will use the [dplyr](http://dplyr.tidyverse.org/index.html) and [readr](http://readr.tidyverse.org/index.html) packages. The data sets can be downloaded from [here](https://github.com/rsquaredacademy/datasets) and the codes from [here](https://gist.github.com/aravindhebbali/55c4f40476028c09949b73af97bb1619).

```{r dplyr1, message=FALSE}
library(dplyr)
library(readr)
options(tibble.width = Inf)
```

<br>

## Case Study

Let us look at a case study (e-commerce data) and see how we can use dplyr helper functions to answer questions we have about and to modify/transform the underlying data set. You can download the data from [here](https://github.com/rsquaredacademy/datasets/blob/master/web.csv) or import it directly using `read_csv()` from the [readr](readr.tidyverse.org) package.

<br>

### Data

```{r show, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

<br>

### Data Dictionary

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)

<br>

## Data Sanitization

- `distinct()`
- `rename()`

<br>

#### Traffic Sources

```{r dplyr2}
ecom %>%
  distinct(referrer)
```

<br>

#### Device Types

```{r dplyr30}
ecom %>%
  distinct(device)
```

<br>

#### Rename Columns

```{r dplyr3}
ecom %>%
  rename(time_on_site = duration)
```

<br>

## Sampling Data

```{r dplyr4}
ecom %>%
  sample_n(700)
```

```{r dplyr5}
ecom %>%
  group_by(referrer) %>%
  sample_n(100)
```

```{r dplyr6}
ecom %>%
  sample_frac(size = 0.7)
```

```{r dplyr7}
ecom %>%
  group_by(referrer) %>%
  sample_frac(0.3)
```

<br>

## Data Extraction

<br>

#### Extract Device Column

```{r dplyr8}
ecom %>%
  pull(device)
```

<br>

#### Extract First Column

```{r dplyr9}
ecom %>%
  pull(1)
```

<br>

#### Extract Last Column

```{r dplyr10}
ecom %>%
  pull(-1)
```

<br>

#### Extract First 20 Rows

```{r dplyr11}
ecom %>%
  slice(1:20)
```

<br>

#### Extract Last Row

```{r dplyr12}
ecom %>%
  slice(n())
```

<br>

## Data Tabulation

<br>

#### Total Observations in Data

```{r dplyr13}
ecom %>%
  tally()
```

<br>

#### Tabulate Referrers

```{r dplyr14}
ecom %>%
  group_by(referrer) %>%
  tally()
```

<br>

#### Tabulate referrers and bouncers

```{r dplyr15}
ecom %>%
  group_by(referrer, bouncers) %>%
  tally()
```

<br>

#### Tabulate referrers and purchasers

```{r dplyr16}
ecom %>%
  group_by(referrer, purchase) %>%
  tally()
```

<br>

#### Tabulate Referrers & Converts

```{r dplyr25}
ecom %>%
  group_by(referrer, purchase) %>%
  tally() %>%
  filter(purchase == 'true')
```

<br>

#### Count

```{r dplyr17}
ecom %>%
  count(referrer, purchase)
```

<br>

## Sort Data

<br>

#### Top 2 referrers by orders

```{r dplyr18}
ecom %>%
  count(referrer, purchase) %>%
  filter(purchase == 'true') %>%
  arrange(desc(n)) %>%
  top_n(n = 2)
```

<br>

## Between

```{r dplyr19}
ecom %>%
  pull(n_pages) %>%
  between(5, 15) 
```

<br>

## Case When

```{r dplyr20}
mtcars %>%
  select(mpg, disp, cyl, gear, carb) %>%
  mutate(
    type = case_when(
      disp > 200 ~ 'High',
      cyl == 8 ~ 'Eight',
      TRUE ~ 'True'
    )
  )
```

<br>

## Select Observations

<br>

#### Select First Observation

```{r dplyr21}
ecom %>%
  pull(referrer) %>%
  nth(1)

ecom %>%
  pull(referrer) %>%
  first()
```

<br>

#### Select 1000th Observation

```{r dplyr22}
ecom %>%
  pull(referrer) %>%
  nth(1000)
```

<br>

#### Select Last Observation

```{r dplyr23}
ecom %>%
  pull(referrer) %>%
  last()
```