---
title: 'ggplot2: Histogram'
date: '2017-06-06'
slug: ggplot2-histogram
topics:
  - ggplot2
tags:
  - ggplot2
  - visualization
---

### Introduction

This is the eleventh post in the series **Elegant Data Visualization with ggplot2**. In the previous [post](https://blog.aravindhebbali.com/), we learnt to build box plots. In this post, we will learn to

- build histogram
- specify bins
- modify 
    - color
    - fill
    - alpha
    - bin width
    - line type
    - line size

A histogram is a plot that can be used to examine the shape and spread of  continuous data. It looks very similar to a bar graph and can be used to detect outliers and skewness in data. The histogram graphically shows the following:

- center (location) of the data
- spread (dispersion) of the data
- skewness
- outliers
- presence of multiple modes

To construct a histogram, the data is split into intervals called bins. The intervals may or may not be equal sized. For each bin, the number of data points that fall into it are counted (frequency). The Y axis of the histogram represents the frequency and the X axis represents the variable. 


```{r hist, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(readr)
```

### Libraries, Code & Data

We will use the following libraries in this post:

- [readr](https://readr.tidyverse.org/)
- [ggplot2](https://ggplot2.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) and code can be downloaded from [here](https://gist.github.com/rsquaredacademy/674bc30cc1539d735bdc4e6210982d1d).

### Data 

```{r import, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

### Data Dictionary

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)

#### Histogram

In ggplot2, a histogram is created using `geom_histogram()`. In the below example, 
we build the histogram of `n_visit` from the `ecom` data. 

```{r hist2, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit))
```

#### Specify Bins

The default number of bins used by `geom_histogram()` is 30 and may not be always useful.
Let us specify the number of bins using the `bins` argument.

```{r hist4, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7)
```

#### Fill

Use the `fill` argument to modify the background color of the histogram. In the below
example, we set the background color to blue.

```{r hist3, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7, fill = 'blue')
```

#### Alpha

The opacity of the background color of the histogram can be modified 
using the `alpha` argument. It takes values between `0` and `1`.

```{r hist11, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7, fill = 'blue', alpha = 0.3)
```

#### Color

The color of the histogram borders can be modified using the `color` argument.

```{r hist8, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7, fill = 'white', color = 'blue')
```

#### Summing up what we have learnt so far..

```{r hist9, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 7, fill = 'blue', color = 'white')
```

#### Bin Width

The width of the bins can be specified using the `binwidth` argument. In the 
below example, we set the bin width to be 2.

```{r hist5, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), binwidth = 2, fill = 'blue', color = 'black')
```

#### Border 

The type and size of the histogram border/line can be modified using
`linetype` and `size`.  In the following examples, we set `linetype` 
to 3 and `size` to 1.25.

###### Line Type

```{r hist6, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 5, fill = 'white', 
    color = 'blue', linetype = 3)
```

###### Line Size

```{r hist10, fig.align='center'}
ggplot(ecom) +
  geom_histogram(aes(n_visit), bins = 5, fill = 'white', 
    color = 'blue', size = 1.25)
```

### Summary

In this post, we learnt to:

- build histogram
- specify bins
- modify 
    - color
    - fill
    - alpha
    - bin width
    - line type
    - line size

### Up Next..

In the next [post](https://blog.aravindhebbali.com/), we will learn to modify the axes of a plot.

