---
title: 'ggplot2: Scatter Plots'
date: '2016-04-04'
slug: ggplot2-scatter-plots
topics:
  - ggplot2
tags:
  - ggplot2
  - visualization
---

### Introduction

This is the fifth post in the series **Elegant Data Visualization with ggplot2**. In the previous [post](https://blog.aravindhebbali.com/), we learnt about text annotations. In this post, we will build scatter plots by applying everything we have learnt so far.

- build scatter plots
- modify point
    - color
    - fill
    - alpha
    - shape
    - size
- fit regression line

```{r scat1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(readr)
```

### Libraries, Code & Data

We will use the following libraries in this post:

- [readr](https://readr.tidyverse.org/)
- [ggplot2](https://ggplot2.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) and code can be downloaded from [here]().

### Basic Plot

Let us build a simple scatter plot using the `mtcars` data to examine the relationship 
between `disp` (diaplacement) and `mpg` (miles per gallon).

```{r scat2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg))
```

#### Jitter

```{r scat22, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), position = 'jitter')
```

#### Aesthetics

Now let us modify the aesthetics of the points. There are two ways:

- map them to variables using the `aes()` function
- or specify values

In the next 4 examples, we will 

- map color to a categorical variable
- map color to a continuous variable
- specify value for color
- specify value for color opacity

###### Map Color to Variable (Categorical)

```{r scat4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl)))
```

###### Map Color to Variable (Continuous)

```{r scat5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = hp))
```

###### Specify Value for Color

```{r scat6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), color = 'blue')
```

###### Specify Value for Alpha

```{r scat7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), color = 'blue', alpha = 0.4)
```

In the next 2 examples, we will 

- map shape to a variable 
- specify value for shape

###### Map Shape to Variable

```{r scat8, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, shape = factor(cyl)))
```

###### Specify Value for Shape

```{r scat9, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), shape = 3)
```

In the next 2 examples, we will 

- map size to a variable 
- specify value for size

###### Map Size to Variable

```{r scat10, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, size = hp), color = 'blue')
```

###### Specify Value for Size

```{r scat11, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg), size = 3)
```

### Regression Line

Most often, after building a scatter plot to examine the relationship between two variables, we fit
a regression line. In this section, we will learn to fit a line to the scatter plot using:

- `geom_smooth()`
- `geom_abline()`

In the below example, we fit a regression line using `geom_smooth()`. It takes two arguments:

- `method` 
- `se`

The `method` argument is used to specify the model type i.e. `lm` or `loess`. 

```{r aes12, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE)
```

###### Regression Line - Loess Method

```{r aes14, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point() +
  geom_smooth(method = 'loess', se = FALSE)
```

###### Regression Line - Conf. Interval

The `se` argument  takes logical values i.e. `TRUE` or `FALSE`. If set to `TRUE`, the confidence band for the regression line is drawn.

```{r aes13, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point() +
  geom_smooth(method = 'lm', se = TRUE)
```

###### Fit Line - Intercept & Slope

We can fit a regression line using `geom_abline()` as well. It takes 2 arguments:

- `slope`
- `intercept`

We can get the `slope` and `intercept` by building a model using `lm()`.

```{r aes17, fig.align='center'}
ggplot(mtcars, aes(disp, mpg)) +
  geom_point() +
  geom_abline(slope = 29.59985, intercept = -0.04122)
```

### Summary

In this post, we learnt to:

- build scatter plots
- map aesthetics to variables
- modify axis and legend
- fit regression line

### Up Next..

In the next [post](https://blog.aravindhebbali.com/), we will learn to build line charts.



