---
title: 'ggplot2: Quick Tour'
date: '2017-01-22'
slug: ggplot2-quick-tour
topics:
  - ggplot2
tags:
  - ggplot2
  - visualization
---

### Introduction

This is the second post in the series **Elegant Data Visualization with ggplot2**. In the previous [post](https://blog.aravindhebbali.com/), we understood the concept of grammar of graphics and even built a bar plot step by step while exploring the different components of a plot/chart. In this post, we will 
learn to quickly build a set of plots that are routinely used to explore data.

```{r gg1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
```

### Libraries, Code & Data

We will use the following libraries in this post:

- [readr](https://readr.tidyverse.org/)
- [ggplot2](https://ggplot2.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) 
and code can be downloaded from [here](https://gist.github.com/rsquaredacademy/3b25fa07c60d44d561819d2c6ab77978).

### qplot()

Through out this post, we will use `qplot()` to build all the plots. It is a shortcut for generating
plots and is useful for quick exploration of data.

#### Scatter Plot

Let us build a scatter plot using `qplot()`. The first two arguments specify the data to be
represented on the `X` and `Y` axis. The third argument specifies the name of the data to be 
used.

```{r gg2, fig.align='center', fig.height=5, fig.width=6}
qplot(disp, mpg, data = mtcars)
```

The default geom used to represent the data is the `point`. We can specify other 
geoms using the `geom` argument. In the below example, the data is represented by
both `point` and `line`.

```{r gg3, fig.align='center', fig.height=5, fig.width=6}
qplot(disp, mpg, data = mtcars, geom = c('point', 'line'))
```

Let us map the color of the points to the variable `cyl`. Since `cyl` is
a categorical variable, we use `factor()` to indicate the same.

```{r gg4, fig.align='center', fig.height=5, fig.width=6}
qplot(disp, mpg, data = mtcars, color = factor(cyl))
```

We can map the shape of the geom to a variable as well. Let 
us map shape to `cyl`.

```{r gg5, fig.align='center', fig.height=5, fig.width=6}
qplot(disp, mpg, data = mtcars, shape = factor(cyl))
```

The size of the geom too can be mapped to a variable. Let us map
size to `qsec`. Keep in mind that shape should be mapped to a 
continuous variable.

```{r gg6, fig.align='center', fig.height=5, fig.width=6}
qplot(disp, mpg, data = mtcars, size = qsec)
```

#### Bar Plot

To build a bar plot, we will specify the geom as `bar`. 

```{r gg7, fig.align='center', fig.height=5, fig.width=6}
qplot(factor(cyl), data = mtcars, geom = c('bar'))
```

We can create a stacked bar plot by mapping `fill` to a variable. In the below
example, we map fill to `am`.

```{r gg8, fig.align='center', fig.height=5, fig.width=6}
qplot(factor(cyl), data = mtcars, geom = c('bar'), fill = factor(am))
```

#### Box Plot

Set geom to `boxplot` to create a box plot. 

```{r gg9, fig.align='center', fig.height=5, fig.width=6}
qplot(factor(cyl), mpg, data = mtcars, geom = c('boxplot'))
```

The below example shows how to generate a box plot when we are not comparing it across groups.

```{r gg10, fig.align='center', fig.height=5, fig.width=6}
qplot(factor(1), mpg, data = mtcars, geom = c('boxplot'))
```

#### Line Chart

Line charts can be created by specifying the geom as `line`.

```{r gg12, fig.align='center', fig.height=5, fig.width=6}
qplot(x = date, y = unemploy, data = economics, geom = c('line'))
```

It is possible to modify the appearance of the line plot. In the below example, 
we change to color of the line to red.

```{r gg13, fig.align='center', fig.height=5, fig.width=6}
qplot(x = date, y = unemploy, data = economics, geom = c('line'),
      color = 'red')
```

#### Histogram

To create a histogram, use the `bins` argument and specify the number of bins for the histogram.

```{r gg14, fig.align='center', fig.height=5, fig.width=6}
qplot(mpg, data = mtcars, bins = 5)
```

### Summary

In this post, we learnt to quickly create plots using the `qplot()` function. While useful, it has limitations and can be used only to quickly visualize data. 

### Up Next..

In the next [post](https://blog.arsquaredacademy.github.io/blog/post/ggplot2-introduction-to-geoms/), we will create the same set of plots but using **geoms**. 

