---
title: 'ggplot2: Line Graphs'
date: '2017-04-19'
slug: ggplot2-line-graphs
topics:
  - ggplot2
tags:
  - ggplot2
  - visualization
---

### Introduction

This is the 8th post in the series **Elegant Data Visualization with ggplot2**. In the previous [post](https://blog.aravindhebbali.com/), we learnt to build scatter plots. In this post, we will learn to 

- build 
    - simple line chart
    - grouped line chart
- map aesthetics to variables
- modify line
    - color
    - type
    - size


```{r line, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

### Libraries, Code & Data

We will use the following libraries in this post:

- [readr](https://readr.tidyverse.org/)
- [ggplot2](https://ggplot2.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) and code can be downloaded from [here](https://gist.github.com/rsquaredacademy/84b3204eee81d6e804f10a73900809b5).


#### Case Study

We will use a data set related to GDP growth rate. You can download it from [here](https://github.com/rsquaredacademy/datasets/blob/master/gdp.csv). It contains GDP (Gross Domestic Product) growth data for the BRICS (Brazil, Russia, India, China, South Africa) for the years 2000 to 2005.

#### Data

```{r import2, eval=TRUE, message=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
gdp
```

#### Line Chart

```{r line100, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line()
```

#### Line Color

The color of the line can be modified using the `color` argument in `geom_lin()`.

```{r line1, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(color = 'blue')
```

#### Line Type

- 0 : blank
- 1 : solid
- 2 : dashed
- 3 : dotted
- 4 : dotdash
- 5 : longdash
- 6 : twodash


#### Line Type

The type of line can be specified using `linetype` argument.

```{r line2, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(linetype = 2)
```

#### Line Type (Dashed)

```{r line3, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(linetype = 'dashed')
```

#### Line Size

The width of the line can be specified using the `size` argument.

```{r line4, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(size = 2)
```

#### Modify Data

Now let us map the aesthetics to the variables. The data used in the above example cannot be used as we need a variable with country names. We will use `gather()` function from the `tidyr` package to reshape the data.

```{r line5, fig.align='center'}
gdp2 <- 
  gdp %>% 
  select(year, growth, india, china) %>% 
  gather(key = country, value = gdp, -year)
gdp2
```

#### Grouped Line Chart

To create multiple lines, we can use the `group` argument and map it to a categorical variable. 
In the below example, we want to visualize the trend in GDP of different countries. Instead of
using `geom_line()` multiple times, we map `group` argument to the variable `country` and we can
visualize the GDP trend for all the countries at once.

```{r line6, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line()
```

We can map aesthetics such as:

- color
- line type 
- and line width

to categorical variables. In the next 3 examples, we map color, line type and line width to the
`country` variable.

###### Map Color to Country

```{r line7, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(color = country))
```

###### Map Line Type to Country

```{r line8, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(linetype = country))
```

###### Map Line Width to Country

```{r line9, fig.align='center'}
ggplot(gdp2, aes(year, gdp, group = country)) +
  geom_line(aes(size = country))
```

### Summary

In this post, we learnt to:

- build 
    - simple line chart
    - grouped line chart
- map aesthetics to variables
- modify line
    - color
    - type
    - size

### Up Next..

In the next [post](https://blog.aravindhebbali.com/), we will learn to build bar plots.

