---
title: 'ggplot2: Box Plots'
date: '2017-05-23'
slug: ggplot2-box-plots
topics:
  - ggplot2
tags:
  - ggplot2
  - visualization
---

### Introduction

This is the 9th post in the series **Elegant Data Visualization with ggplot2**. In the previous [post](https://blog.aravindhebbali.com/), we learnt how to build bar charts. In this post, we will 

- build box plots
- modify box 
    - color
    - fill
    - alpha
    - line size
    - line type
- modify outlier 
    - color
    - shape 
    - size 
    - alpha

```{r box, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(readr)
```

### Libraries, Code & Data

We will use the following libraries in this post:

- [readr](https://readr.tidyverse.org/)
- [ggplot2](https://ggplot2.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) and code can be downloaded from [here](https://gist.github.com/rsquaredacademy/246091b512a6c006e68374e2d24caf7c).

### Data

```{r import, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

### Data Dictionary

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)

#### Univariate Box Plot

Let us create a univariate box plot i.e. we are not comparing the distribuion
of the variable across groups. In `geom_boxplot()`, we must map the `x` aesthetic 
to a variable else it will return an error. In order to create the box plot, we will
assing the value `factor(1)` to the `x` aesthetic and the variable whose distribution
we are examining to the `y` aesthetic.

```{r box2, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(1), y = n_visit))
```

#### Box Plot

If we are comparing the distribution of a variable across groups, we can map the
`x` aesthetic to a categorical variable. In the below example, we are comparing the
distribution of `n_visit` across different device types.

```{r box3, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = n_visit))
```

#### Notch

If we want to test whether the medians of the different groups differ, we can use the `notch` argument and set it to `TRUE`. A notch is drawn in each side of the boxes and if the notches of the plots do not overlap, it is strong evidence that the medians differ.

```{r box5, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = n_visit),
               notch = TRUE) 
```

#### Outliers

The box plot is useful in detecting outliers in the data. In this section, we will learn 
to modify the appearance of the outlier using:

- `outlier.color`
- `outlier.shape`
- `outlier.size`
- `outlier.opacity`

In the next 4 examples, we will modify the appearance of the outlier.

###### Outlier Color

```{r box7, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.color = 'red')
```

###### Outlier Shape

```{r box8, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.shape = 23)
```

###### Outlier Size

```{r box10, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration), outlier.size = 3) +
  expand_limits(y = c(0, 1100))
```

###### Outlier Alpha

```{r box11, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               outlier.color = 'blue', outlier.alpha = 0.3) +
  expand_limits(y = c(0, 1100))
```

### Box Aesthetics

In this section, we will learn to modify the appearance of the box.

#### Background Color

The background color of the box can be modified using the `fill` argument. We can
either map it to variables or specify values for each box in the plot.

###### Specify Values for Fill

```{r box12, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               fill = c('blue', 'red', 'green')) 
```

###### Map Fill to Variable

```{r box13, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration,
               fill = factor(device))) 
```

#### Opacity

The opacity of the background color can be modified using the `alpha` argument.
It takes values between `0` and `1`.

```{r box14, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               fill = 'blue', alpha = 0.3) 
```

#### Box Border

The color of the box line or border can be modified using the `color` argument. Like `fill`,
we can map it to a categorical variable or specify values for each box.

```{r box15, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               color = c('blue', 'red', 'green')) 
```

###### Map Color to Variables

```{r box16, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration,
               color = factor(device))) 
```

#### Line Width & Type

The width and type of the box line can be modified using the `size` and
`linetype` arguments.

###### Specify Values for Line Width

```{r box17, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               size = 1.5) 
```

###### Specify Values for Line Type

```{r box18, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = duration),
               linetype = 2) 
```


#### Horizontal Box Plot

Use `coord_flip()` to create a horizontal box plot.

```{r box4, fig.align='center'}
ggplot(ecom) +
  geom_boxplot(aes(x = factor(device), y = n_visit)) +
  coord_flip()
```


### Summary

In this post, we learnt to:

- build box plots
- modify outlier color, shape, size etc.
- modify box color
- modify box line color, size and line type

### Up Next..

In the next [post](https://blog.aravindhebbali.com/), we will learn to build histograms.

