---
title: 'ggplot2: Introduction to geoms'
date: '2017-02-11'
slug: ggplot2-introduction-to-geoms
topics:
  - ggplot2
tags:
  - ggplot2
  - visualization
---

### Introduction

This is the third post in the series **Elegant Data Visualization with ggplot2**. In the previous [post](https://blog.aravindhebbali.com/), we learnt how to create plots using the `qplot()` function. In this post, we will create some of the most routinely used plots to explore data using the `geom_*` functions. 

```{r geoms1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(tibble)
```

### Libraries, Code & Data

We will use the following libraries in this post:

- [readr](https://readr.tidyverse.org/)
- [ggplot2](https://ggplot2.tidyverse.org/)
- [tibble](https://tibble.tidyverse.org/)
- [dplyr](https://dplyr.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) and code can be downloaded from [here](https://gist.github.com/rsquaredacademy/2d0d5d5b60d0ef0f4d1b227c8fb0335d).

### Data

```{r use, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

#### Data Dictionary

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)

### Scatter Plot

A scatter plot displays the relationship between two continuous variables. In ggplot2, we can build a scatter plot using `geom_point()`. Scatterplots can show you visually:

- the strength of the relationship between the variables 
- the direction of the relationship between the variables 
- and whether outliers exist

#### Point

```{r geoms2, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg)) + 
  geom_point()
```

#### Horizontal/Vertical Lines

Add horizontal or vertical lines using:

- `geom_hline()`
- `geom_vline()`

To add a horizontal line, we need to specify the location on the Y axis where 
the line will be drawn. Use `yintercept` to specify the location of the line.

```{r geoms4, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_hline(yintercept = 30) 
```

For the vertical line, we need to specify the location on the X axis using 
`xintercept`.

```{r geoms3, fig.align='center'}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_vline(xintercept = 5) 
```

#### Bar Plot

Bar plots present grouped data with rectangular bars. The bars may represent the frequency of the groups or values. Bar plots can be: 

- horizontal 
- vertical
- grouped
- stacked 
- proportional

Let us build a simple bar plot to visualize the traffic driven by different device types.

```{r geoms8, fig.align='center'}
ggplot(ecom, aes(x = factor(device))) +
  geom_bar()
```

Instead of the frequency of visits, we can visualize the order value driven by the 
different devices. Use the `weight` argument within `aes` to indicate that the bars 
should represent the variable specified and not the frequency.

```{r geoms9, fig.align='center'}
ggplot(ecom, aes(x = factor(device))) +
  geom_bar(aes(weight = order_value))
```

To view the proportion of purchasers and non-purchasers for each device type, we will
map `fill` to `purchase`. The color of the bar represents purchasers and non-purchasers.

```{r geoms10, fig.align='center'}
ggplot(ecom, aes(x = factor(device))) +
  geom_bar(aes(fill = purchase))
```

#### Columns

In some instances, we do not have access to the raw data and are provided summaries or 
transformed data. In the below example, we have data that summarizes the visits from 
each device type. Such data can be visualized using `geom_col()`. 

```{r geoms12, fig.align='center', fig.height=4}
device <- c('laptop', 'mobile', 'tablet')
visits <- c(30000, 12000, 5000)
traffic <- tibble::tibble(device, visits)
ggplot(traffic, aes(x = device, y = visits)) +
  geom_col(fill = 'blue') 
```

#### Boxplot

Box plots can be used to:

- examine the distribution of a variable  
- detect outliers, boxplots are very handy

`geom_boxplot()` is used to create box plots. 

```{r geoms13, fig.align='center'}
ggplot(ecom, aes(x = factor(device), y = n_pages)) +
  geom_boxplot()
```

We must specify both the `x` and `y` aesthetic. If you are not comparing the distribution
across groups, use the below method to generate the box plot.

```{r geoms13a, fig.align='center'}
ggplot(ecom, aes(x = factor(1), y = n_pages)) +
  geom_boxplot()
```

#### Histogram

Histograms are used to examine:

- distribution of a continuous variable
- skewness and kurtosis

We can create a histogram using `geom_histogram()`. Only the `x` aesthetic needs to be supplied.

```{r geoms15, fig.align='center'}
ggplot(ecom, aes(x = duration)) +
  geom_histogram()
```

The default number of bins used is 30 which may not be helpful always. Use `bins` argument 
to specify the appropriate number of bins for the histogram.

```{r geoms16, fig.align='center'}
ggplot(ecom, aes(x = duration)) +
  geom_histogram(bins = 5)
```

#### Line

Line charts are used to examine trends over time.

#### Data

```{r import2, eval=TRUE, message=FALSE}
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
gdp
```

Use `geom_line()` to create line plots.

```{r geomline1, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line()
```

The following can be modified to improve the appearance of the line:

- `color`
- `size`
- `linetype` 

In the below example, we modify the color and type of the line.

```{r geomline2, fig.align='center'}
ggplot(gdp, aes(year, india)) +
  geom_line(color = 'blue', linetype = 'dashed')
```


#### Label

Labels can be added to identify data points using `geom_label()`.

```{r geoms20, fig.align='center'}
ggplot(mtcars, aes(disp, mpg, label = rownames(mtcars))) +
  geom_label()
```

#### Text

Use `geom_text()` to add text to the plot. We can modify the size of the text 
and ensure that they do not overlap using `check_overlap` argument.

```{r geoms19, fig.align='center'}
ggplot(mtcars, aes(disp, mpg, label = rownames(mtcars))) +
  geom_text(check_overlap = TRUE, size = 2)
```

In the below example, we: 

- map the color of the text to `cyl` variable 
- nudge the text to avoid overlapping with the points
- and change the horizontal justification, size and angle of text

```{r geoms21, fig.align='center'}
ggplot(mtcars, aes(x = disp, y = mpg, label = rownames(mtcars))) +
  geom_point() + 
  geom_text(aes(color = cyl), hjust = 0, nudge_x = 0.05, 
            size = 2, angle = 45)
```

### Summary 

In this post, we learnt to build different types of plots using `geom_*` instead of `qplot()`.

### Up Next..

In the next [post](https://blog.aravindhebbali.com), we will learn about aesthetics.

