---
title: 'ggplot2: Bar Plots'
date: '2017-05-07'
slug: ggplot2-bar-plots
topics:
  - ggplot2
tags:
  - ggplot2
  - visualization
---

### Introduction

This is the ninth post in the series **Elegant Data Visualization with ggplot2**. In the previous [post](https://blog.aravindhebbali.com/), we learnt to build line charts. In this post, we will learn to:

- build 
    - simple bar plot
    - stacked bar plot
    - grouped bar plot
    - proportional bar plot
- map aesthetics to variables
- specify values for 
    - bar color
    - bar line color
    - bar line type
    - bar line size

```{r bar, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(readr)
```

### Libraries, Code & Data

We will use the following libraries in this post:

- [readr](https://readr.tidyverse.org/)
- [ggplot2](https://ggplot2.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) and code can be downloaded from [here](https://gist.github.com/rsquaredacademy/096329693fa1f9313771d4e259cce1ec).

### Data

```{r import2, eval=TRUE, message=FALSE}
ecom <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/web.csv')
ecom
```

### Data Dictionary

- id: row id
- referrer: referrer website/search engine
- os: operating system
- browser: browser
- device: device used to visit the website
- n_pages: number of pages visited
- duration: time spent on the website (in seconds)
- repeat: frequency of visits
- country: country of origin
- purchase: whether visitor purchased
- order_value: order value of visitor (in dollars)

### Aesthetics

- `fill`
- `color`
- `linetype`
- `size`
- `position`


#### Simple Bar Plot

Let us start by building a simple bar plot using `geom_bar()`. We will look at the
distribution of devices that drive traffic to a fictional website.

```{r bar2, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)))
```

#### Bar Color

Use the `fill` argument to modify the color of the bars.

```{r bar3, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)),
           fill = c('red', 'blue', 'green'))
```

#### Stacked Bar Plot

To build a stacked bar plot, map the `fill` argument to another categorical variable. In 
the below example, we map `fill` to `referrer`.

```{r bar7, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)))
```

#### Grouped Bar Plot

Use the `position` argument to create a grouped bar plot. Assign the value `dodge` 
to `position` argument. Instead of stacking the bars on top of one another, the bar 
are placed next to each other. 

```{r bar8, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)), position = 'dodge')
```

#### Proportional Bar Plot

Proportional bar plots can be created by assigning the value `fill` to 
the `position` argument. In a proportional bar plot, the height of all 
the bars is same.

```{r bar10, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(device, fill = factor(referrer)), position = 'fill')
```

#### Horizontal Bar Plot

Use `coord_flip()` to create a horizontal bar plot.

```{r bar9, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device), fill = factor(referrer))) +
  coord_flip()
```

In the next 3 examples, we will learn to modify the following properties of the bar lines:

- line color
- line type
- and line size

#### Bar Line Color

The color of the bar lines can be specified using the `color` argument.

```{r bar4, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white',
           color = c('red', 'blue', 'green'))
```

#### Bar Line Type

To modify the line type of the bars, use the `linetype` argument.

```{r bar5, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white', 
          color = 'black', linetype = 2)
```

#### Bar Line Size

The `size` argument can be used to modify the size of the bar lines.

```{r bar6, fig.align='center'}
ggplot(ecom) +
  geom_bar(aes(factor(device)), fill = 'white', 
           color = 'black', size = 2)
```

### Summary

In this post, we learnt to:

- build 
    - simple bar plot
    - stacked bar plot
    - grouped bar plot
    - proportional bar plot
- map aesthetics to variables
- specify values for 
    - bar color
    - bar line color
    - bar line type
    - bar line size


### Up Next..

In the next [post](https://blog.aravindhebbali.com/), we will learn to build box plots.

