---
title: Guides - Axes
date: '2017-06-26'
slug: guides-axes
topics:
  - data-visualization
tags:
  - ggplot2
---

### Introduction

This is the twelfth post in the series **Elegant Data Visualization with ggplot2**. In the previous [post](https://blog.aravindhebbali.com/), we learnt to build histograms. Now that we have learnt to build different plots, let us look at different ways to modify the axis. Along the way, we will also explore the `scale_*()` family of functions.

Modify X and Y axis

- title
- labels
- limits
- breaks
- position

In this module, we will learn how to modify the X and Y axis using the following functions:

- Continuous Axis
    - `scale_x_continuous()`
    - `scale_y_continuous()`

- Discrete Axis
    - `scale_x_discrete()`
    - `scale_y_discrete()`

### Libraries, Code & Data

We will use the following libraries in this post:

- [readr](https://readr.tidyverse.org/)
- [ggplot2](https://ggplot2.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) and code can be downloaded from [here](https://gist.github.com/rsquaredacademy/096bc745c18a9ba47b99260978189920).

```{r load, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(readr)
```

### Continuous Axis

`scale_x_continuous()` and `scale_y_continuous()` take the following arguments:

- name
- limits
- breaks
- labels
- position

Let us create a scatter plot using `geom_point()` to examine the relationship between `disp` (displacement) and
`mpg` (miles per gallon) from the mtcars data set.

```{r guide2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg))
```

###### Axis Label

```{r guide3, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(name = "Displacement")
```

###### Axis Limits

```{r guide4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(limits = c(0, 600))
```

###### Axis Breaks

```{r guide5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(breaks = c(0, 150, 300, 450, 600))
```

###### Axis Tick Labels

```{r guide6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(breaks = c(0, 150, 300, 450, 600),
    labels = c('0', '150', '300', '450', '600'))
```

###### Axis Position

```{r guide7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  scale_x_continuous(position = 'top')
```

#### X Axis 

```{r guide8, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg)) +
  scale_x_continuous(name = "Displacement", limits = c(0, 600),
                     breaks = c(0, 150, 300, 450, 600), position = 'top',
                     labels = c('0', '150', '300', '450', '600'))
```

#### Y Axis

```{r guide9, fig.align='center'}
ggplot(mtcars) + geom_point(aes(disp, mpg)) +
  scale_y_continuous(name = "Miles Per Gallon", limits = c(0, 45),
                     breaks = c(0, 15, 30, 45), position = 'right',
                     labels = c('0', '15', '30', '45'))
```

### Discrete Axis

`scale_x_discrete()` and `scale_y_discrete()` take the following arguments:

- name
- labels
- breaks
- position

###### Axis Label

```{r guide11, fig.align='center'}
ggplot(mtcars) +
  geom_bar(aes(factor(cyl))) +
  scale_x_discrete(name = "Number of Cylinders")
```

###### Axis Tick Labels

```{r guide12, fig.align='center'}
ggplot(mtcars) +
  geom_bar(aes(factor(cyl))) +
  scale_x_discrete(labels = c("4" = "Four", "6" = "Six", "8" = "Eight"))
```

###### Axis Breaks

```{r guide13, fig.align='center'}
ggplot(mtcars) +
  geom_bar(aes(factor(cyl))) +
  scale_x_discrete(breaks = c("4", "6", "8"))
```

###### Axis Position

```{r guide14, fig.align='center'}
ggplot(mtcars) +
  geom_bar(aes(factor(cyl))) +
  scale_x_discrete(position = 'bottom')
```

###### X Axis - Discrete 

```{r guide15, fig.align='center'}
ggplot(mtcars) + geom_bar(aes(factor(cyl))) +
  scale_x_discrete(name = "Number of Cylinders",
    labels = c("4" = "Four", "6" = "Six", "8" = "Eight"),
    breaks = c("4", "6", "8"), position = "bottom")
```

### Summary

In this post, we learnt to modify the following component of the X and Y axis:

- title
- labels
- limits
- breaks
- position

### Up Next..

In the next [post](https://blog.aravindhebbali.com/), we will learn to modify the legend when `color` is mapped to a variable.

