---
title: 'ggplot2: Text Annotations'
date: '2017-03-17'
slug: ggplot2-text-annotations
topics:
  - ggplot2
tags:
  - ggplot2
  - visualization
---

### Introduction

This is the sixth post in the series **Elegant Data Visualization with ggplot2**. In the previous [post](https://blog.aravindhebbali.com/), we learnt to modify the axis and plot labels. In this post, we will learn to add text to the plots.

- add custom text 
- modify color
- modify size
- modify fontface
- modify angle

### Libraries, Code & Data

We will use the following libraries in this post:

- [readr](https://readr.tidyverse.org/)
- [ggplot2](https://ggplot2.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) and code can be downloaded from [here](https://gist.github.com/rsquaredacademy/567094325b5e134b5b4459677206f363).

#### Annotate

We will use the `annotate()` function to add custom text to the plots. You can use the `annotate()` function to add rectangles/segments/pointrange as well but our focus will be on adding text. Let us start with a simple scatter plot.

```{r ann1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(readr)
```

Use `annotate()` to add custom text to a plot. It takes the following arguments:

- `geom`     : specify text
- `x`        : x axis location
- `y`        : y axis location
- `label`    : custom text 
- `color`    : color of text
- `size`     : size of text
- `fontface` : fontface of text
- `angle`    : angle of text

##### Add Text

```{r ann2, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text')
```

##### Modify Color

```{r ann4, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', color = 'red')
```

##### Specify Size

```{r ann5, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', size = 6)
```

##### Change Font

```{r ann6, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', fontface = 'bold')
```

##### New Angle

```{r ann7, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text', angle = 25)
```

##### Sum it all up...

```{r ann3, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg)) +
  annotate('text', x = 200, y = 30, label = 'Sample Text',
           color = 'red', size = 6, fontface = 'bold', angle = 25)
```

#### Summary

In this post, we learnt to add custom text to a plot and modify:

- color
- size
- font
- angle 

#### Up Next..

In the next [post](https://blog.aravindhebbali.com/), we will build scatter plots.

