---
title: 'ggplot2: Legends - Part 1'
author: Aravind Hebbali
twitterImg: /img/gg_leg_part_1.png
description: "ggplot2 - Quick Tour"
date: '2018-02-12'
slug: legend-part-1
categories:
  - data-visualization
tags:
  - ggplot2
---

### Introduction

This is the 13th post in the series **Elegant Data Visualization with 
ggplot2**. In the previos post, we learnt how to modify the axis of plots. In 
this post, we will focus on modifying the legends of plots when the aesthetics 
are mapped to variables. If the aesthetics are mapped to variables, ggplot2 
automatically creates legends wherever applicable and you may want to modify the 
appearance of legends. In this post, we will learn to modify the legends 
when `color` is mapped to categorical variables. Specifically, we will learn
to modify:

- title
- breaks
- limits
- labels
- values

```{r leg1, echo=FALSE, eval=TRUE, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

<br>

### Libraries, Code & Data

We will use the following libraries in this post:

- [readr](http://readr.tidyverse.org/)
- [ggplot2](http://ggplot2.tidyverse.org/)

All the data sets used in this post can be found [here](https://github.com/rsquaredacademy/datasets) 
and code can be downloaded from [here](https://gist.github.com/rsquaredacademy/f099b954fa8f5a84cd8e5a2a031f91db).

<br>

### Basic Plot

Let us start with a scatter plot examining the relationship between displacement
and miles per gallon from the mtcars data. We will map the color of the points
to the `cyl` variable. 

```{r leg15, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl)))
```

<br>

### Values

```{r leg17, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"))
```

<br>

### Title

The title can be modified using the `name` argument in `scale_color_manual()`. 

```{r leg16, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(name = "Cylinders", 
    values = c("red", "blue", "green"))
```

<br>

### Limits

Modify the data to be displayed.

```{r leg18, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    limits = c(5, 6, 8))
```

<br>

### Labels

```{r leg19, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    labels = c('Four', 'Six', 'Eight'))
```

<br>

### Breaks

Breaks to be displayed in the legend.

```{r leg20, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(values = c("red", "blue", "green"),
    breaks = c(4, 8))
```

<br>

### Putting it all together...

```{r leg21, fig.align='center'}
ggplot(mtcars) +
  geom_point(aes(disp, mpg, color = factor(cyl))) +
  scale_color_manual(name = "Cylinders", values = c("red", "blue", "green"),
    labels = c('Four', 'Six', 'Eight'), limits = c(4, 6, 8), breaks = c(4, 6, 8))
```

<br>

### Summary

In this post, we learnt to modify the following aspects of legends: 

- title
- breaks
- limits
- labels
- values

<br>

### Up Next..

In the next post, we will learn how to modify legend when `fill` is mapped to variables.

<br>