---
title: 'Working with Dates in R '
author: Aravind Hebbali
date: '2017-12-25'
slug: working-with-dates-in-r
categories:
  - data wrangling
tags:
  - lubridate
  - date
  - time
---

## Introduction



 In this post, we will learn to work with date/time data in R using [lubridate]( http://lubridate.tidyverse.org/), an R package that makes it easy to work with dates and time. Let us begin by installing and loading the pacakge.
 
## Libraries, Code & Data

We will use [lubridate](http://lubridate.tidyverse.org/index.html), [dplyr](http://dplyr.tidyverse.org/index.html), [magrittr](http://magrittr.tidyverse.org/index.html) and [readr](http://readr.tidyverse.org/index.html) packages. The data sets can be downloaded from [here](https://github.com/rsquaredacademy/datasets) and the codes from [here](https://gist.github.com/aravindhebbali/7758b86c2bc13ff1e5d88d9d1c204f8c).

```{r lub1, message=FALSE}
library(lubridate)
library(dplyr)
library(magrittr)
library(readr)
```

## Quic Intro

#### Origin

```{r origin}
lubridate::origin
```

#### Current Date/Time

```{r today}
now()
today()
am(now())  
pm(now())
```

## Case Study

We will use the functions in the lubridate package to answer a few 
questions we have about the transact data.

- extract date, month and year from Due
- compute the number of days to settle invoice
- compute days over due
- check if due year is a leap year
- check when due day in february is 29, whether it is a leap year
- how many invoices were settled within due date
- how many invoices are due in each quarter
- what is the average duration between invoice date and payment date


### Data

```{r import, eval=FALSE}
transact <- read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/transact.csv')
```

```{r show, echo=FALSE, eval=TRUE, message=FALSE}
transact <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/transact.csv')
transact
```

### Data Dictionary

The data set has 3 columns. All the dates are in the format (yyyy-mm-dd).

- Invoice: invoice date 
- Due: due date
- Payment: payment date

## Extract Date, Month & Year from Due Date

```{r lub18}
this_day <- as_date('2017-03-23')
day(this_day)
month(this_day)
year(this_day)
```

```{r lub2}
transact %>%
  mutate(
    due_day = day(Due),
    due_month = month(Due),
    due_year = year(Due)
  )
```

## Compute days to settle invoice

```{r lub19}
course_start <- as_date('2017-04-12')
course_end <- as_date('2017-04-21')
course_duration <- course_end - course_start
course_duration
```

```{r lub3}
transact %>%
  mutate(
    days_to_pay = Payment - Invoice
  )
```

## compute days over due

```{r lub4}
transact %>%
  mutate(
    delay = Payment - Due
  )
```

## Is due year a leap year?

```{r lub5}
transact %>%
  mutate(
    due_year = year(Due),
    is_leap = leap_year(due_year)
  )
```

## If due day is February 29, is it a leap year?

```{r lub6}
transact %>%
  mutate(
    due_day = day(Due),
    due_month = month(Due),
    due_year = year(Due),
    is_leap = leap_year(due_year)
  ) %>%
  select(-(Invoice), -(Payment)) %>%
  filter(due_month == 2 & due_day == 29)
```

## Shift Date

```{r lab40}
course_start + days(2)
course_start + weeks(1)
course_start + years(1)
```

## Interval

```{r lub10}
interval(course_start, course_end)
```

## Shift Interval

```{r lab50}
course_interval <- interval(course_start, course_end)
course_interval %>%
  int_shift(by = days(1))
course_interval %>%
  int_shift(by = weeks(1))
course_interval %>%
  int_shift(by = years(1))
```

## Intervals Overlap

```{r lub60}
vacation_start <- as_date('2017-04-19')
vacation_end <- as_date('2017-04-25')
vacation_interval <- interval(vacation_start, vacation_end)
int_overlaps(course_interval, vacation_interval)
```

## how many invoices were settled within due date?

```{r lub7}
transact %>%
  mutate(
    inv_due_interval = interval(Invoice, Due),
    due_next = Due + days(1),
    due_pay_interval = interval(due_next, Payment),
    overlaps = int_overlaps(inv_due_interval, due_pay_interval)
  ) %>%
  select(Invoice, Due, Payment, overlaps)
```

````{r lub12}
# using int_shift
transact %>%
  mutate(
    inv_due_interval = interval(Invoice, Due),
    due_pay_interval = interval(Due, Payment),  
    due_pay_next = int_shift(due_pay_interval, by = days(1)),
    overlaps = int_overlaps(inv_due_interval, due_pay_next)
  ) %>%
  select(Invoice, Due, Payment, overlaps)
```

## Within

```{r lub30}
conference <- as_date('2017-04-15')
conference %within% interval(course_start, course_end)
```

#### how many invoices were settled within due date?

```{r lub13}
transact %>%
  mutate(
    inv_due_interval = interval(Invoice, Due),
    overlaps = Payment %within% inv_due_interval
  ) %>%
  select(Due, Payment, overlaps)
```

## Quarter

```{r lub20}
course_start
course_start %>%
  quarter()
course_start %>%
  quarter(with_year = TRUE)
```

```{r lub8}
transact %>%
  mutate(
    Quarter = quarter(Due)
  )
```

```{r lub14}
transact %>%
  mutate(
    Quarter = quarter(Due, with_year = TRUE)
  )
```

## Duration

```{r lub15}
course_duration %>%
  as.numeric() %>%
  duration(units = 'days')
```

#### Case Study

```{r lub11}
interval(course_start, course_end) / dseconds()
interval(course_start, course_end) / dminutes()
interval(course_start, course_end) / dhours()
interval(course_start, course_end) / dweeks()
interval(course_start, course_end) / dyears()
```


```{r lub16}
interval(course_start, course_end) %>%
  time_length(unit = "seconds")
interval(course_start, course_end) %>%
  time_length(unit = "minutes")
interval(course_start, course_end) %>%
  time_length(unit = "hours")
```

```{r lub17}
interval(course_start, course_end) %>%
  as.period(unit = "seconds")
interval(course_start, course_end) %>%
  as.period(unit = "minutes")
interval(course_start, course_end) %>%
  as.period(unit = "hours")
```